---
title: "p value extraction for plotting"
author: "andrew hall"
date: '2022-06-09'
output:
  pdf_document: default
  html_document: default
---

\section{Introduction}

This markdown document uses the mock data from the R workshop folder; you can read in your own data and replace the variable names to complete anova & Tukey,  extract significance from the test output, and incorporate the significance letters on a bar plot.

You can use this script to generate a report of your data, as well as to create a figure and export to .png


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(multcompView)
library(RColorBrewer)
library(gt)
library(knitr)
```

\subsection{Output of Statistical Tests}

```{r read in data and conduct statistical test, echo=FALSE}
##Read in dummy data from R workshop
data <- read.table('starch_data.txt', header=TRUE, sep='\t')
fit<-aov(starch~genotype, data=data)
tuk <- TukeyHSD(fit, "genotype", ordered=TRUE)

summary.aov(fit)
plot(tuk)

print(tuk)
```
\subsection{Worked up Mean and SD Data with Significance Labels}

```{r work up data for plotting, echo=FALSE}

##Extract p-values from the Tukey output using multcompView
pvalues<-extract_p(tuk)
pvalueslist<-pvalues$genotype
##Turn p-values vector into a matrix
pvaluesmat<-vec2mat(pvalueslist, sep="-")
##Use multcompLetters command to determine significance for each genotype
###You can adjust which p-value threshold you want to use!
sigletters<-multcompLetters(pvaluesmat, compare="<", threshold=0.05, Letters=c(letters, LETTERS, "."), reversed=FALSE)
##Make significance letters into a data frame

sigletters<-sigletters$Letters
sigletters<-as.data.frame(sigletters)
##Make row names into a normal column
rows<-row.names(sigletters)
row.names(sigletters)=NULL
sigletters<-cbind(genotype=rows, sigletters)
sigletters<-sigletters%>%
  arrange(genotype)##this is important!
sigvec<-c(sigletters$sigletters)

##Make a new data frame with means and SDs for creating bar graph
datasum<-data%>%
  group_by(genotype)%>%
  summarise_at(vars(starch), list(mean=mean, sd=sd))%>%
  arrange(genotype)%>%
  cbind(sigletters=sigvec)%>%
  ##Reorder your groups in custom order (ie put WT first)
  mutate(genotype=fct_relevel(genotype, "WT", "MutA", "MutB", "MutC", "MutD"))%>%
  as_tibble()

datasum%>%
  gt()%>%
  tab_header(title="Summary Data")
```
\subsection{Generation of Bar Plots}
```{r make plot, echo=FALSE}

##Now you have everything needed to make a ggplot

starch_plot<-ggplot()+
  ##add columns with the mean data
  geom_col(data=datasum, width=0.5, aes(x=genotype, y=mean, fill=genotype))+
  ##you can customize the colors of each bar
  scale_fill_manual(values=c("WT"="grey",
                             "MutA"="maroon",
                             "MutB"="maroon",
                             "MutC"="maroon",
                             "MutD"="maroon"
  ))+
  ##add in all the data points from the raw data
  geom_point(data=data, aes(x=genotype, y=starch))+
  ##add error bars
  geom_errorbar(data=datasum, width=0.1, aes(ymin=mean-sd, ymax=mean+sd, x=genotype))+
  ##add significance labels
  geom_text(data=datasum, aes(x=genotype, y=(mean+sd+3), label=sigletters))+
  ylab("Starch Content")+
  xlab("Genotype")+
  theme(
    legend.position="none",
    plot.background = element_rect(color="white"),
    panel.background=element_blank(),
    axis.line = element_line(color="black")
  )
starch_plot

#optionally, you can export your plot to .png using the following command (uncomment to use):

#ggsave("plotname.png", plot=starch_plot, dpi=1000)


```